---
title: "Customising plots with ggplot2"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  out.width  = "100%"
)
```

```{r setup}
library(ggvariant)
library(ggplot2)

vcf_file <- system.file("extdata", "example.vcf", package = "ggvariant")
variants <- read_vcf(vcf_file)
```

Every `ggvariant` plot function returns a standard `ggplot` object (not a
subclass, not a wrapped list), so any `ggplot2` layer, scale, theme, or
annotation composes onto it exactly as it would onto a plot you built by
hand.

## Adding layers

```{r add-layers}
plot_lollipop(variants, gene = "KRAS") +
  labs(subtitle = "KRAS mutations in cohort X") +
  theme(legend.position = "bottom")
```

## Overriding a scale

`ggvariant` sets a fill or colour scale internally, so replacing it means
adding a new `scale_*` call after the plot is built -- `ggplot2` uses the
last matching scale in the layer stack.

```{r override-scale}
plot_consequence_summary(variants) +
  scale_fill_brewer(palette = "Set2", name = "Consequence")
```

## Building on `theme_ggvariant()`

`theme_ggvariant()` is exported, so you can start from it rather than
`theme_minimal()` when building a custom plot from scratch, or layer
further `theme()` tweaks onto a `ggvariant` plot's existing theme.

```{r theme-tweak}
plot_variant_spectrum(variants) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title  = element_text(colour = "grey20")
  )
```

## Combining with `gv_palette()`

The built-in palettes are exported as plain named character vectors, so
they work equally well outside a `ggvariant` plot function -- for example,
when building a custom `ggplot2` call directly against a `gvf` object.

```{r gv-palette}
ggplot(variants, aes(x = consequence, fill = consequence)) +
  geom_bar() +
  scale_fill_manual(values = gv_palette("consequence"), guide = "none") +
  coord_flip() +
  theme_ggvariant()
```
