---
title: "Gallery"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  out.width  = "100%"
)
```

```{r setup}
library(ggvariant)

vcf_file <- system.file("extdata", "example.vcf", package = "ggvariant")
variants <- read_vcf(vcf_file)
```

This gallery collects one example of every `ggvariant` plot type against the
bundled example VCF. See `vignette("ggvariant")` for a narrated walkthrough,
and the function reference for the full argument list of each plot.

## Lollipop plot

```{r lollipop}
plot_lollipop(variants, gene = "TP53")
```

## Consequence summary

```{r consequence-summary}
plot_consequence_summary(variants, group_by = "gene", top_n = 6)
```

## Mutational spectrum

```{r spectrum}
plot_variant_spectrum(variants, facet_by_sample = TRUE)
```

## Oncoprint / waterfall plot

`plot_oncoprint()` and `plot_waterfall()` are the same function under the two
names used by different communities — "oncoprint" in the `ComplexHeatmap`/
cBioPortal vocabulary, "waterfall plot" in the `GenVisR`/`maftools`
vocabulary. A waterfall plot and an oncoprint are the same visualisation:
`plot_waterfall()` and `plot_oncoprint()` are aliases for one function and
produce identical output — use whichever name your field uses.

The bundled example VCF has only two samples, too few to show the cascade
staircase this plot is built around. The cohort below is illustrative
synthetic data generated just for this article (not the package's example
data), sized like a real cohort so the staircase and the annotation track
are both visible, with a clinical `stage` track added via `annotation`:

```{r oncoprint-cohort}
set.seed(42)

genes <- c("TP53", "KRAS", "PIK3CA", "EGFR", "BRCA1",
           "BRCA2", "APC", "PTEN", "ARID1A", "SMAD4")
gene_freq <- c(TP53 = 0.85, KRAS = 0.55, PIK3CA = 0.40, EGFR = 0.30,
               BRCA1 = 0.22, BRCA2 = 0.18, APC = 0.14, PTEN = 0.10,
               ARID1A = 0.07, SMAD4 = 0.05)
samples <- sprintf("COHORT_S%02d", 1:28)
consequences <- c("missense_variant", "stop_gained",
                   "frameshift_variant", "synonymous_variant")

cohort <- do.call(rbind, lapply(genes, function(g) {
  n_mut <- max(1, round(gene_freq[[g]] * length(samples)))
  data.frame(
    chrom = "chr1", pos = sample(1e6:2e6, n_mut),
    ref = "A", alt = "T", gene = g,
    sample = sample(samples, n_mut),
    consequence = sample(consequences, n_mut, replace = TRUE),
    stringsAsFactors = FALSE
  )
}))
cohort_variants <- coerce_variants(cohort)

clinical_stage <- data.frame(
  sample = samples,
  stage  = sample(c("I", "II", "III", "IV"), length(samples), replace = TRUE)
)

plot_waterfall(cohort_variants, top_n = 10, annotation = clinical_stage)
```

## Tumour mutational burden

```{r tmb}
plot_tmb(variants)
```
