---
title: "Gallery"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  out.width  = "100%"
)
```

```{r setup}
library(ggvariant)

vcf_file <- system.file("extdata", "example.vcf", package = "ggvariant")
variants <- read_vcf(vcf_file)
```

This gallery collects one example of every `ggvariant` plot type against the
bundled example VCF. See `vignette("ggvariant")` for a narrated walkthrough,
and the function reference for the full argument list of each plot.

## Lollipop plot

```{r lollipop}
plot_lollipop(variants, gene = "TP53")
```

## Consequence summary

```{r consequence-summary}
plot_consequence_summary(variants, group_by = "gene", top_n = 6)
```

## Mutational spectrum

```{r spectrum}
plot_variant_spectrum(variants, facet_by_sample = TRUE)
```

## Oncoprint / waterfall plot

`plot_oncoprint()` and `plot_waterfall()` are the same function under the two
names used by different communities.

```{r oncoprint}
plot_oncoprint(variants, top_n = 6)
```

With a clinical annotation track:

```{r oncoprint-annotation}
clinical <- data.frame(
  sample = c("TUMOR_S1", "TUMOR_S2"),
  stage  = c("III", "IV")
)
plot_oncoprint(variants, top_n = 6, annotation = clinical)
```

## Tumour mutational burden

```{r tmb}
plot_tmb(variants)
```
