---
title: "Interactive plots with plotly"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  out.width  = "100%"
)
has_plotly <- requireNamespace("plotly", quietly = TRUE)
```

Every `ggvariant` plot function accepts `interactive = TRUE`, which routes
the plot through [plotly](https://plotly.com/r/) instead of returning a
static `ggplot` object. This is useful for sharing a plot with collaborators
who don't use R: a plotly widget can be saved as a standalone HTML file and
opened in any browser.

`plotly` is a Suggests dependency, not a hard dependency of `ggvariant` --
the chunks on this page only run when `plotly` is installed on the machine
building the site.

```{r setup}
library(ggvariant)

vcf_file <- system.file("extdata", "example.vcf", package = "ggvariant")
variants <- read_vcf(vcf_file)
```

## Interactive lollipop plot

Hovering over a point shows the position, amino-acid change, consequence,
and sample.

```{r lollipop-interactive, eval = has_plotly}
plot_lollipop(variants, gene = "TP53", interactive = TRUE)
```

## Interactive oncoprint

```{r oncoprint-interactive, eval = has_plotly}
plot_oncoprint(variants, top_n = 6, interactive = TRUE)
```

## Saving a widget for sharing

```{r save-widget, eval = FALSE}
p <- plot_lollipop(variants, gene = "TP53", interactive = TRUE)
htmlwidgets::saveWidget(p, "TP53_lollipop.html")
```

`htmlwidgets::saveWidget()` is not a formal dependency of `ggvariant` --
install it separately (`install.packages("htmlwidgets")`) if you want to
export a standalone file this way.
